Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance

### 2019-04-18 Manuscript/manuscript.Rmd
manuscript edits

Lastly, we investigated which genomic annotations were most predictive of TAD boundary formation. Looking at heatmaps of variable importances provided by the random forest classifiers implemented on SMOTE re-sampled data, with distance-type predictors, we found that known transcription factors associated with the hypothesized loop extrusion model [@fudenberg2016formation; @alipour2012self] were give high predictive importance in the model for both cell lines (Figure 6 A-B). These tended to include the SMC3ab, ZNF143, CTCF, and RAD21 transcription factor binding sites and were found to be ranked consistly across resolution. Other annotations known to be associated with gene regulation and affect structural properties of chromatin, such as the YY1 and MAZ transcription factors, as well DNAse hypersensitive sites, regions of heterochromatin, and the H3k27 trimethylation and H3K4me1 modifications were less consistently ranked across resolutions for either cell line. Moreover, the ranking profiles of genomic annotations were found to be more similar between the 10 kb and 25 kb resolutions, as well as between the 50 kb and 100 kb resoluions.

>>>>

We then investigated which genomic annotations were most predictive of TAD boundary formation, across resolution for random forest models with the best given set of data-level characteristics determined above (SMOTE re-sampled with distance type predictors). We found that specific transcription factor binding sites, including SMC3ab, ZNF143, CTCF, and RAD21, were given consistently high rankings of predictive importance in the models built on the GM12878 cell line (Table 1). These are known factors associated with the hypothesized loop extrusion model [@fudenberg2016formation; @alipour2012self] and were found to be ranked consistently high on the K562 cell line as well, with similar rankings across resolutions (Supplementary Table 6). Other annotations known to be associated with gene regulation and affect structural properties of chromatin, such as the YY1  transcription factor, as well DNAse hypersensitive sites, regions of heterochromatin, and the H3k27 trimethylation histone modifications were ranked high on the 10 kb resolution, but were less consistently ranked at lower resolutions for both cell lines. <!--Moreover, the ranking profiles of genomic annotations for both cell lines were found to be more similar between the 10 kb and 25 kb resolutions, as well as between the 50 kb and 100 kb resoluions (this is proved when looking at a clustered heatmap).-->
